DATA 621 Homework #4

### 2019-11-09 hw4/HW4.Rmd
Update HW4

```{r}
df <- read.csv("./data/insurance_training_data.csv") %>%
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
  select(-INDEX)
evaluation <- read.csv("./data/insurance-evaluation-data.csv") %>%
  select(-INDEX)

@@

Specifically, there `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable are missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, but only `r df %>% filter(is.na(YOJ),is.na(CAR_AGE)) %>% nrow()` observations missing both of these varables. So, if imputation is not viable here, we will lose as many as 488 observations out of the 8161 in our training set (or about 6%).

@@

There are some variables that are currently factors that are dollar values that need to be transformed into numeric variables.  We will do this to both the `df` and `evaluation` data frames.

@@

df <- fix_data_types(df)
evaluation <- fix_data_types(evaluation)
```

@@

### Fix Missing Values
There are missing values.  We will fill in the missing variables using KNN.

```{r, null_prefix = TRUE}
# set.seed(42)
# knn <- df %>%
#   select(-TARGET_FLAG, -TARGET_AMT, is.numeric()) %>%
#   #na.omit() %>%
#   knnImputation()
# 
# for(var in names(df)){
#   impute_me <- is.na(df[[var]])
#   if (nrow(df[impute_me,]) > 0){
#     print(paste("Fixing the", nrow(df[impute_me,]), "missing values in", var, "(df)"))
#     #df[impute_me, var] <- knn[impute_me, var] 
#   }
# }
```

@@

>>>>

```{r}
df <- read.csv("./data/insurance_training_data.csv")
evaluation <- read.csv("./data/insurance-evaluation-data.csv")

@@

@@

There are some variables that are currently factors that are dollar values that need to be transformed into numeric variables.  We will do this to both the `df` and `evaluation` data frames.  There are also some invalid data that will be changed to NAs.

@@

na_bad_values <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(CAR_AGE = ifelse(CAR_AGE < 0, NA, CAR_AGE))%>%
    ungroup()
}

df <- df %>%
  fix_data_types() %>%
  na_bad_values()
evaluation <- evaluation %>%
  fix_data_types() %>%
  na_bad_values()
```

@@

### Fix Missing Values

Specifically, there `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable are missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, `r df %>% filter(is.na(AGE)) %>% nrow() ` observations where the variable `AGE` is missing, `r df %>% filter(is.na(INCOME)) %>% nrow() ` observations where the variable `INCOME` is missing, and `r df %>% filter(is.na(HOME_VAL)) %>% nrow() ` observations where the variable `HOME_VAL` is missing.  There are `r nrow(df) - nrow(na.omit(df))`, or `r round(((nrow(df) - nrow(na.omit(df))) / nrow(df)) * 100)`% of the observations missing varables.

@@

There are `r nrow(train[TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.  We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.

```{r}
set.seed(42)
minority <- nrow(train[train$TARGET_FLAG == 1,])
majority <- nrow(train[train$TARGET_FLAG == 0,])
diff <- majority - minority
minority_index <- train[train$TARGET_FLAG == 1,]$INDEX
over_sample_train <- data.frame(INDEX = sample(minority_index, diff, T)) %>%
  merge(train, .) %>%
  bind_rows(train)
```
The over sampled data frame has `r nrow(over_sample_train)` records, and as you can see in the following figure is now balanced:

```{r, echo=FALSE}
over_sample_train %>% 
  ggplot(aes(x=TARGET_FLAG, fill=TARGET_FLAG)) +
  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
  theme_light() +
  theme(legend.position = "none") +
  labs(x="TARGET_FLAG", y="# Observations")
```
